Responses to the reviews of ACL rolling review paper \#162: ``Latent Group Dropout for Multilingual and Multidomain Machine Translation''

### 2022-01-14 2022/acl/acl_latex.tex

\begin{abstract}
  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalities of the tasks at hand, while smaller parts are reserved to model the peculiarities of a language or a domain. In adapter-based approaches, these strategies are hardcoded in the network architecture, independent of the similarities between tasks. In this work, we propose a new method to better take advantage of these similarities, using a latent-variable model. We also develop new techniques to train this model end-to-end and report experimental results showing that the learned patterns are both meaningful and yield improved translation performance without any increase of the model size. 
\end{abstract}

@@

\textbf{Language similarity} The methods developed by \citep{sen19multilingual,kong21multilingual} use language proximity to design parameter sharing strategies. The authors propose a multi-decoder model sharing the same encoder among languages and routing languages in different families to different decoders. These approaches share the same interest in expressing the proximity between tasks in the selection of task-specific parameters as our approach. However, our method learn the selection from a latent commonality in data instead of using a predefined selection such as "One language family per decoder" in \citep{kong21multilingual}.

\textbf{Language-specific sub-networks}. \done\todo{}\citet{Frankle19Lottery,liu18rethinking} study techniques to identify the most important parameters for the current task, so that masking the less important parameters during training does not hurt performance. \citet{lin21learning} adapts this idea for multilingual NMT, trying to identify language dependent subsets of parameters by pruning a fine-tuned model. Our approach also aims to map sub-networks to tasks: we do so by masking the output of each layer, rather than masking parameters. Furthermore, \citet{lin21learning} computes the masks via a heuristic selection; while our approach learns the masks with variational techniques.

>>>>

\begin{abstract}
  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalities of the tasks at hand, while smaller parts are reserved to model the peculiarities of a language or a domain. In adapter-based approaches, these strategies are hardcoded in the network architecture, independent of the similarities between tasks. In this work, we propose a new method to better take advantage of these similarities, using a latent-variable model. We also develop new techniques to train this model end to end and report multiple experimental results that show the learned patterns are both meaningful and yield improved translation performance without any increase of the model size. 
\end{abstract}

@@

\textbf{Language similarity} The methods developed by \citep{sen19multilingual,kong21multilingual} use language proximity to design parameter sharing strategies. The authors propose a multi-decoder model sharing the same encoder among languages and routing languages in different families to different decoders. These approaches share the same interest in expressing the proximity between tasks in the selection of task-specific parameters as our approach. However, our method learn the selection from a latent commonality in data instead of using a predefined selection such as "One language family per decoder" in \cite{kong21multilingual}.

\textbf{Language-specific sub-networks}. \todo{}\citet{Frankle19Lottery,liu18rethinking} proposed different techniques to identify most important parameters for the current task so that when training the model while masking other parameters the performance of the model does not degrade. \citet{lin21learning} hypothesized that among parameters of a multilingual NMT model, there exists a subset of parameters more biased to a specific language. The authors proposed to locate these subsets by pruning the parameters of each fine-tuned model. Our approach also aims to assign a sub-network to each task by masking the output of each layer rather than masking the parameters. Furthermore, \citet{lin21learning} computes the masks via a heuristic selection while our approach learn the masks via a variational technique.

### 2022-01-14 2022/acl/responses.tex

However, we would also like to outline the fact that our contribution, in this paper, is *not* to outperform the best published state-of-the-art results for multilingual / multidomain MT. Instead, our goal is to present a novel, mathematically sound, methodology to automatically learn shared nodes in multi-task (multi-domain, multi-lingual) systems and to show (a) that these new patterns yield performance that are better that alternative approaches for sharing / unsharing subpart of the computation graph accross tasks and also compare favorably in terms of computational cost; (b) that these patterns actually capture similarities between tasks. We have therefore planned our experiments, comparisons and analyses insofar as they were relevant to illustrate these aspects. This has notably motivated our choice the adapter model, which has been used in many prior studies, as a reasonable point of comparison. We however respectfully disagree on the fact that unpublished papers from mid 2021 cited in the reviews should be the ``baseline'' for reporting experiments in november 2021.

@@

\subsubsection*{Rebuttal}
{\color{blue}%
We thank the reviewer for pointing additional relevant work. We cite and discuss it in the revised version.\Todo{Check this}

For the sake of the clarification of our approach, we would like to point out several misunderstandings in the summary made by Reviewer iMf4:
\begin{itemize}
\item Our approach does not aim to find task-specific and task-agnostic \emph{parameters} in the neural network: there is no notion of domain-specific or domain-agnostic parameters in the approach we present. The objective of our algorithm is to optimize the assignment \emph{of a sub-group of "nodes"}, at the output of each intermediate layer, to each task.
\item Our approach does not mask the sub-parts of each weight matrix, but instead it masks a fixed number of sub-groups of elements of the output of each intermediate layer.
\end{itemize}
We have tried to improve our presentation of our method in the revised version of the paper to avoid further misunderstandings.\todo{Change presentation, give line numbers}

We would like to answer the other concerns of the reviewier as follows. \todo{CLAIM : baseline + parameter sharing + resultats CDR}
\begin{itemize} 
\item Regarding baselines, we generally refer to the rebuttal to the meta review. Note that according to your suggestion, we have added new multidomain experiments with a heuristic definition of blocks. See the updated Table $3$.\Todo{Add table}
\item We have now included an analysis on the effect of $k$ on the performance in the revised version (see new section 4.4)
\item The multilingual experiments use two groups of $8$ languages, which are all low-resourced; this might lower the significance of the improvement as observed in the paper of Lin et al. 21 (only +0.7 for low resource languages). Our method significantly improves the less resources languages among these groups.
\item As suggested, we are now using heatmaps instead of Table 7.
\item Statistical confidence scores are also reported. \todo{Add significancy tests}
\end{itemize}
}

@@

We agree that a mixture of experts model would also provide yet another possible alternative baseline. Note however that mixture of experts are much costly to train, much more than our method, as we would need to fully train one model for each expert; furthermore, a mixture of expert would not bring us any insight regarding the benefits of sharing / unsharing patterns for related domains. Therefore, we do not see that the outcome of such comparison would allow us to revise or amend or main claims.
}
\section*{Reviewer 3koi}

@@

\begin{itemize}
\item We initialize the probabilistic distribution of the latent variables as uniform distribution. This is clarified in the revised text in Section$3.1.4$.\done\Todo{Check clarified which line}
\item The revised version includes a discussion regarding $k$ and how it affects the performance of our approach (section 4.4).
\item As reported in many works using the Gumbel reparameterization trick, the effect of $\tau$ is relatively insignificant. This is discussed in the revised version in section 3.1.3. \done\Todo{which section}Therefore, we focused more on selecting $k$. \Todo{Add references}
\item Because we would like to have the same \emph{total amount of training iterations between Adapter-based and LaMGD}, we extend the training of LaMGD to 300K. The convergence of the standard Transformer is achieved before 200K as its validation curve became flat near the 200K-th iteration. These has been clarified in the paper in section 3.1.1. \done\Todo{Which section ?} 
\item The same answer applies to the reviewer's concern regarding the multilingual experiments.
\item As pointed out\done\todo{Where ?} in Table~5, significant gains are observed in several very low-resourced languages, not globally. This might be explained by the fact that low-resourced languages benefit most from the sharing nodes with a related well-resourced languages, and are also less affected by the inference of unrelated languages.
\end{itemize}

Our claim on the advantage in computational cost wrt. Adapters relies on the inference stage while there are no extra parameters added to the standard Transformer compared to Adapter-based methods. At the same time, the training time will be more than usual because of the joint optimization of the network's organization and the parameters. \todo{Not so great. Can we say more ? There is a nice blog post on FLOPS here: \url{%
https://medium.com/@dzmitrybahdanau/the-flops-calculus-of-language-model-training-3b19c1f025e4
}%
}
}%

>>>>

However, we would also like to outline the fact that our contribution, in this paper, is not to outperform the best published state-of-the-art results for multilingual / multidomain MT. Instead, our goal is to present a novel, mathematically sound, methodology to automatically learn shared nodes in multi-task (multi-domain, multi-lingual) systems and to show (a) that these new patterns yield performance that are better that alternative approaches for sharing / unsharing subpart of the computation graph accross tasks and also compare favorably in terms of computational cost; (b) that these patterns actually capture similarities between tasks. We have therefore planned our experiments, comparisons and analyses insofar as they were relevant to illustrate these aspects. This has notably motivated our choice the adapter model, which has been used in many prior studies, as a reasonable point of comparison. We however respectfully disagree on the fact that unpublished papers from mid 2021 cited in the reviews should be the ``baseline'' for reporting experiments in november 2021.

@@

\subsubsection*{Rebuttal}
{\color{blue}%
We thank the reviewer for suggesting additional relevant work. We cite and discuss it in the revised version in section 5.\done\Todo{Check this}

For the sake of the clarification of our approach, we would like to point out several misunderstandings in the summary made by Reviewer iMf4:
\begin{itemize}
\item Our approach does not aim to find task-specific and task-agnostic \emph{parameters} in the neural network: there is no notion of domain-specific or domain-agnostic parameters in the approach we present. The objective of our algorithm is to optimize the assignment \emph{of a sub-group of "nodes"}, at the output of each intermediate layer, to each task.
\item Our approach does not mask the sub-parts of each weight matrix, but instead it masks a fixed number of sub-groups of elements of the output of each intermediate layer.
\end{itemize}
We have tried to improve our presentation of our method in the revised version of the paper to avoid further misunderstandings (see lines 92-95, lines 115-119).\done\todo{Change presentation, give line numbers}

We would like to answer the other concerns of the reviewer as follows. \done\todo{CLAIM : baseline + parameter sharing + resultats CDR}
\begin{itemize} 
\item Regarding baselines, we generally refer to the rebuttal to the meta review. Note that according to your suggestion, we have added new multidomain experiments with a heuristic definition of blocks. See the updated Table $3$.\done\Todo{Add table}
\item We have now included an analysis on the effect of $k$ on the performance in the revised version (see new section 4.4)
\item The multilingual experiments use two groups of $8$ languages, which are all low-resourced; this might lower the significance of the improvement as observed in the paper of Lin et al. 21 (only +0.7 for low resource languages). Our method significantly improves the less resources languages among these groups.
\item As suggested, we are now using heatmaps instead of Table 7.
\item Statistical confidence scores are also reported in Section 3.2. \done\todo{Add significancy tests}
\end{itemize}
}

@@

We agree that a mixture of experts model would also provide yet another possible alternative baseline. Note however that mixture of experts are much costly to train, much more than our method, as we would need add new sets of parameters for each expert. Furthermore, we do not see that the outcome of such comparison would allow us to revise or amend or main claims.
}
\section*{Reviewer 3koi}

@@

\begin{itemize}
\item We initialize the probabilistic distribution of the latent variables as a uniform distribution. This is clarified in the revised text in Section~3.1.4.\done\Todo{Check clarified which line}
\item The revised version includes a discussion regarding $k$ and how it affects the performance of our approach (section 4.4).
\item As reported in many works using the Gumbel reparameterization trick, the effect of $\tau$ is relatively insignificant. This is discussed in the revised version in Section 3.1.3. \done\Todo{which section}Therefore, we focused more on selecting $k$. \done\Todo{Add references}
\item Because we would like to have the same \emph{total amount of training iterations between Adapter-based and LaMGD}, we extend the training of LaMGD to 300K. The convergence of the standard Transformer is achieved before 200K as its validation curve became flat near the 200K-th iteration. These has been clarified in the paper in Section~3.1.1. \done\Todo{Which section ?} 
\item The same answer applies to the reviewer's concern regarding the multilingual experiments.
\item As pointed out\done\todo{Where ?} in Table~5, significant gains are observed in several very low-resourced languages, not globally. We have checked that these differences were significant. This might be explained by the fact that low-resourced languages benefit most from the sharing nodes with a related well-resourced languages, and are also less affected by the inference of unrelated languages.
\end{itemize}

Our claim on the advantage in computational cost wrt. Adapters relies on the inference stage while there are no extra parameters added to the standard Transformer compared to Adapter-based methods. At the same time, the training time will be more than usual because of the joint optimization of the network's organization and the parameters. \done\todo{Not so great. Can we say more ? There is a nice blog post on FLOPS here: \url{%
https://medium.com/@dzmitrybahdanau/the-flops-calculus-of-language-model-training-3b19c1f025e4
}%
}
}%
